High-Dimensional Regularized Discriminant Analysis

### 2016-01-11 reports/paper-hdrda.Rnw
Updated results. Added summaries for new data sets

We compared the \emph{HDRDA} classifer to the five competing classifiers on four
benchmark gene-expression microarray data sets. We evaluated the classification
performance of each classifier by randomly partitioning the data set under
consideration such that $2/3$ of the observations were allocated as training
data and the remaining $1/3$ of the observations were allocated as a test data
set. After training each classifier on the training data, we classified the test
data sets and computed the proportion of mislabeled test observations to
estimate the classification error rate for each classifier. Repeating this
process 250 times, we computed the average of the error-rate estimates for each
classifier. Here, we provide a concise description of each high-dimensional
data set examined in our classification study.

\subsubsection{\cite{Burczynski:2006ik} Data Set}

\cite{Burczynski:2006ik} acquired peripheral blood mononuclear cells (PBMC)
through hybridization to microarrays from 127 individuals resulting in 22,283
sequences. Of the 127 individuals, 42 were healthy, 59 had Crohn's disease (CD),
and 26 had ulcerative colitis (UC). The goal of \cite{Burczynski:2006ik} was to
improve accuracy in the discrimination of Inflammatory Bowel Disease (IBD) using
the PBMC-based gene expression signature of a patient.

@@

@@

Similar to \cite{Witten:2011kc}, we report the average test error rates obtained
over 250 random training-test partitions in Table \ref{tab:microarray-results}
along with standard deviations of the test error rates in parentheses. The
\emph{HDRDA} and Guo classifiers were superior in classification performance for
the majority of the simulations with the popular Guo classifier providing
superior classification performance for the Burczysnki data set. Both
\emph{HDRDA} methods yielded comparable results.

<<classification_results, results='asis'>>=
table_caption <- 'The average of the test error rates obtained on gene-expression data sets over 250 random training-test partitions. Standard deviations of the test error rates are given in the parentheses.'

>>>>

We compared the \emph{HDRDA} classifer to the five competing classifiers on four
benchmark gene-expression microarray data sets. We evaluated the classification
accuracy of each classifier by randomly partitioning the data set under
consideration such that $2/3$ of the observations were allocated as training
data and the remaining $1/3$ of the observations were allocated as a test data
set. To expedite the computational runtime, we reduced the training data to the
top 1000 variables by employing the variable-selection method proposed by
\cite{Dudoit:2002ev}. We then reduced the test data set to the same 1000
variables. After training each classifier on the training data, we classified
the test data sets and computed the proportion of mislabeled test observations
to estimate the classification error rate for each classifier. Repeating this
process 100 times, we computed the average of the error-rate estimates for each
classifier. Here, we provide a concise description of each high-dimensional data
set examined in our classification study.

\subsubsection{\cite{Chiaretti:2004gq} Data Set}

\cite{Chiaretti:2004gq} measured the gene-expression profiles for 128
individuals with acute lymphoblastic leukaemia (ALL) using Affymetrix human
95Av2 arrays. Following \cite{Xu:2009fl}, we restricted the data set to $K = 2$
classes such that $n_1 = 74$ observations were without cytogenetic abnormalities
and $n_2 = 37$ observations have a detected BCR/ABL gene. The robust multichip
average normalization method was applied to all 12,625 gene expression levels.

\subsubsection{\cite{Chowdary:2006kf} Data Set}

\cite{Chowdary:2006kf} investigated 52 matched pairs of tissues from colon and
breast tumors using Affymetrix U133A arrays and ribonucleic-acid (RNA)
amplification. Each tissue pair was gathered from the same patient and
consisted of a snap-frozen tissue and a tissue suspended in a RNAlater
preservative. Overall, 31 breast-cancer and 21 colon-cancer pairs were gathered,
resulting in $K = 2$ classes with $n_1 = 62$ and $n_2 = 42$. A purpose of the
study was to determine whether the disease state could be identified using
22,283 gene-expression profiles.

@@

\subsubsection{\cite{Tian:2003ht} Data Set}

\cite{Tian:2003ht} investigated the purified plasma cells from the bone marrow
of control patients along with patients with newly diagnosed multiple
myeloma. Expression profiles for 12,2625 genes were obtained via Affymetrix
U95Av2 microarrays. The plasma cells were subjected to biochemical and
immunohistochemical analyses to identify molecular determinants of osteolytic
lesions. For 36 multiple-myloma patients, focal bone lesions could not be
detected by magnetic resonance imaging (MRI), whereas MRI was used to detect
such lesions in 137 patients.

@@

Similar to \cite{Witten:2011kc}, we report the average test error rates obtained
over 100 random training-test partitions in Table \ref{tab:microarray-results}
along with standard deviations of the test error rates in parentheses. The
\emph{HDRDA} and Guo classifiers were superior in classification performance for
the majority of the simulations with the popular Guo classifier providing
superior classification performance for the Burczysnki data set. Both
\emph{HDRDA} methods yielded comparable results.

<<classification_results, results='asis'>>=
table_caption <- 'The average of the test error rates obtained on gene-expression data sets over 100 random training-test partitions. Standard deviations of the test error rates are given in the parentheses.'
